---
title: "Land cover data"
author: "Samuel R.P-J. Ross"
date: "25/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(ggplot2)
require(readr)
require(tibble)
require(tidyr)
require(scales)
require(dplyr)
require(patchwork)

setwd("~/Desktop/Research/In progress/2021 Hannah project/Data")
Site_data<-read_csv(file = 'longform_Site_data.csv',col_names = T)

```

Our land cover data has buffer zones at 3 scales (100m, 400m, and 1000m). I expect that birds are most likely affected at the 1000m scale because of their dispersal ability, but the 100m scale also represents (local) fine-scale land cover differences that might affect habitat use at the local scale. So we'll separate out our land cover data to these two different scales: 

```{r}

# get only local 100m scale
Land_local <- Site_data[Site_data$Buffer %in% 100,]

# get only regional 1000m scale
Land_km <- Site_data[Site_data$Buffer %in% 1000,]

```

Start by visualising our sites in space: 

```{r}

plot(Land_local$xcoord_EPSG_3035,Land_local$ycoord_EPSG_3035)

```

Yep, they look like transects! Just by eyeballing this, I suspect we'll need to account for spatial autocorrelation by fitting models with spatial error structure. 

Next let's have a look at some histograms to see how much variation there is in different land cover types among our sites:

```{r}

# proportion of urban area
hist(Land_local$P_Urban,breaks = 10)
hist(Land_km$P_Urban,breaks = 10)

# proportion of agriculture
hist(Land_local$P_Agriculture,breaks = 10)
hist(Land_km$P_Agriculture,breaks = 10)

# proportion of agriculture (including agriculture with natural patches)
hist(Land_local$`P_Agri_and_semi-agri`,breaks = 10)
hist(Land_km$`P_Agri_and_semi-agri`,breaks = 10)

# proportion of agriculture (including agriculture with natural patches)
hist(Land_local$P_natural,breaks = 10)
hist(Land_km$P_natural,breaks = 10)

```

For proportion of urban area, we see a few sites have some urban area, more so at the larger spatial scale. But these are very much not urban sites - even the highest values are around 35% urban area. 

For proportion of agriculture (excluding agricultural land with significant natural patches), we see a strong dichotomy at the local 100m scale (more suited to a discrete axis, agriculture vs other) but at the 1000m scale there's a nice range of values from 0-1 more suited to a continuous axis (proportion of agriculture).

The same is true for proportion of agriculture including land that is predominantly agriculture but has some natural vegetation too - though in this case, there's a heaver skew towards 1, so the distribution looks less Gaussian. 

When looking only at purely natural land cover (grasses, forests, scrub, marshland etc) we see a similar pattern in the opposite direction (skewed towards 0).

Now let's also take a quick look at our other measures: 

```{r}

# mean tree cover density
hist(Land_local$tcd_mean,breaks = 10)
hist(Land_km$tcd_mean,breaks = 10)

# percentage of small woody features
hist(Land_local$SWF_area_PERCENT_type_TOTAL,breaks = 10)
hist(Land_km$SWF_area_PERCENT_type_TOTAL,breaks = 10)

# mean leaf area index
hist(Land_local$lai_mean,breaks = 10)
hist(Land_km$lai_mean,breaks = 10)

# mean topographic wetness index
hist(Land_local$twi_mean,breaks = 10)
hist(Land_km$twi_mean,breaks = 10)

# mean topographic wetness index
hist(Land_local$elev_m_mean,breaks = 10)
hist(Land_km$elev_m_mean,breaks = 10)

```

Mean tree cover density shows very few trees at the 100m scale, but a better (albeit still heavy tailed) distribution at the 1km scale. 

The % of small woody features (hedges etc) mimics this pattern at the 100m scale - very few SWFs - but looks a bit more normally distributed at the 1km scale.

Leaf area index shows a nice spread, with a hint of a normal distribution at the 1km scale. 

The same is true for the topographic wetness index, representing a nice spread of fast vs slow draining land following rainfall.

Finally, we might also consider elevation, which shows a nice spread of sites at higher vs lower elevations. 

Remember that all of the land cover values will be spatially autocorrelated within these distributions, though! 

~End~